---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Bivariate Alternating Recurrent Event Data Analysis (BivRec)

Alternating recurrent event data arise frequently in biomedical and social sciences where two types of events such as hospital admissions and discharge occur alternatively over time. BivRec implements a collection of non-parametric and semiparametric methods to analyze such data. 

The main functions are:  
+ biv.rec.fit: Use for the estimation of covariate effects on the two alternating event gap times (Xij and Yij) using semiparametric methods. The method options are "Lee.et.al" and "Chang".  
+ biv.rec.np:  Use for the estimation of the joint cumulative distribution funtion (cdf) for the two alternating events gap times (Xij and Yij) as well as the marginal survival function for type I gap times (Xij) and the conditional cdf of the type II gap times (Yij) given an interval of type I gap times (Xij) in a non-parametric fashion.    

The package also provides options to simulate and visualize the data and results of analysis. 

## Installation

BivRec depends on the following system requirements:  
+ Rtools. Download Rtools 35 from https://cran.r-project.org/bin/windows/Rtools/  

Once those requirements are met you can install BivRec from github as follows:
```{r gh-installation, eval = FALSE}
#Installation requires devtools package.
#install.packages("devtools")
library(devtools)
assignInNamespace("version_info", c(devtools:::version_info, 
                                    list("3.5" = list(version_min = "3.3.0", version_max = "99.99.99", 
                                    path = "bin"))), "devtools")
find_rtools()
#Make sure the results of previous command is true
install_github("SandraCastroPearson/BivRec")
```

## Example

This is an example using a simulated data set.

```{r BivRecExample}
# Simulate bivariate alternating recurrent event data
library(BivRec)
set.seed(8822)
biv.rec.data <- biv.rec.sim(nsize=150, beta1=c(0.5,0.5), beta2=c(0,-0.5), tau_c=63, set=1.1)
head(biv.rec.data)

# Plot gap times
biv.rec.plot(formula = id + epi ~ xij + yij, data = biv.rec.data)
```
```{r BivRecExample2, fig.height=4.5}
# Apply the non-parametric method of Huang and Wang (2005) and visualize marginal and conditional results

# To save plots in a pdf file un-comment the following lines of code: 
# pdf("nonparamplots.pdf")
# par(mfrow=c(3,1))
nonpar.result <- biv.rec.np(formula = id + epi + xij + yij + d1 + d2 ~ 1,
           data=biv.rec.data, ai=1, u1 = c(2, 5, 10, 20), u2 = c(1, 5, 10, 15),
           conditional = TRUE, given.interval=c(0, 10), jointplot=TRUE,
           marginalplot = TRUE, condiplot = TRUE)

# To close the pdf file with the saved plots un-comment the following line of code
# dev.off()

head(nonpar.result$joint.cdf)
head(nonpar.result$marginal.survival)
head(nonpar.result$conditional.cdf)

# Apply Lee C, Huang CY, Xu G, Luo X (2017) method using multiple covariates
fit.lee <- biv.rec.fit(formula = id + epi + xij + yij + d1 + d2 ~ a1 + a2,
                data=biv.rec.data, method="Lee.et.al", CI=0.99)
fit.lee$covariate.effects

# To apply Chang (2004) method use method="Chang".
# biv.rec.chang<- biv.rec.fit(formula = id + epi + xij + yij + d1 + d2 ~ a1 + a2, 
# data=biv.rec.data, method="Chang", CI=0.99)
```
