---
title: "Bivariate Alternating Recurrent Event Data Analysis (BivRec)"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

Alternating recurrent event data arise frequently in biomedical and social sciences where two types of events such as hospital admissions and discharge occur alternatively over time. BivRec implements a collection of non-parametric and semiparametric methods to analyze such data. 

The main functions are:  
- bivrecReg: Use for the estimation of covariate effects on the two alternating event gap times (Xij and Yij) using semiparametric methods. The method options are "Lee.et.al" and "Chang".  
- bivrecNP:  Use for the estimation of the joint cumulative distribution funtion (cdf) for the two alternating events gap times (Xij and Yij) as well as the marginal survival function for type I gap times (Xij) and the conditional cdf of the type II gap times (Yij) given an interval of type I gap times (Xij) in a non-parametric fashion.    

The package also provides options to simulate and visualize the data and results of analysis. 

## Installation

BivRec depends on the following system requirements:  
- Rtools. Download Rtools 35 from https://cran.r-project.org/bin/windows/Rtools/  

Once those requirements are met you can install BivRec from github as follows:
```{r gh-installation, eval=FALSE}
#Installation requires devtools package.
#install.packages("devtools")
library(devtools)
#install_github("SandraCastroPearson/BivRec")
```

## Example

This is an example using a simulated data set.

```{r BivRecExample}
# Simulate bivariate alternating recurrent event data
library(BivRec)
set.seed(1234)
simdata <- simulate(nsize=150, beta1=c(0.5,0.5), beta2=c(0,-0.5), tau_c=63, set=1.1)
head(simdata)

# Create a bivrecSurv object
bivrec_data <- with(simdata, bivrecSurv(id, epi, xij, yij, d1, d2))
# Plot gap times
plot(bivrec_data)

# Apply the non-parametric method of Huang and Wang (2005) and visualize marginal and conditional results


nonpar_result <- bivrecNP(response = with(bivrec_data, bivrecSurv(id, epi, xij, yij, d1, d2)),
                      ai=1, u1 = c(2, 5, 10, 20), u2 = c(1, 5, 10, 15),
                      conditional = TRUE, given.interval=c(0, 10))

head(nonpar_result$joint.cdf)
head(nonpar_result$marginal.survival)
head(nonpar_result$conditional.cdf)
plot(nonpar_result)

# To save individual plots in a pdf file un-comment the following line of code: 
# pdf("nonparam_jointcdfplot.pdf")
# plotJoint(nonpar_result)
# dev.off()
# pdf("nonparam_marginalplot.pdf")
# plotMarg(nonpar_result)
# dev.off()
# pdf("nonparam_conditionaplot.pdf")
# plotCond(nonpar_result)
# dev.off()

# Apply Lee C, Huang CY, Xu G, Luo X (2017) method using multiple covariates
fitlee <- bivrecReg(formula = bivrecSurv(id, epi, xij, yij, d1, d2) ~ a1 + a2,
                     data = simdata, method="Lee.et.al")
summary(fitlee)
plot(fitlee)

# To apply Chang (2004) method use method="Chang".
# biv.rec.chang<- bivrecReg(formula = bivrecSurv(id, epi, xij, yij, d1, d2) ~ a1 + a2,
#                     data = simdata, method="Chang")
```
